---
title: "4_phyloseq_data_analysis1"
author: "stinekrye"
date: "2021-09-21"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(phyloseq)
library(tidyverse)
library(readxl)
```


```{r}
load("./data/Old/2_seqtab.nochim.identifier.Rdata")
load("./data/Old/2_taxa.species.rds")
sampleData <- read_excel("./data/Old/1_sampleData.xlsx")
sampleData <- column_to_rownames(sampleData, "ini_month_id")
seqtab.nochim.identifier <- select(seqtab.nochim.identifier, -c("sample_name"))
```


Create phyloseq object
```{r}
asvTable <- otu_table(seqtab.nochim.identifier, taxa_are_rows = FALSE)
sampleData <- sample_data(sampleData)
taxoTable <- tax_table(taxa.species)
phyloData <- phyloseq(asvTable,sampleData, taxoTable)
```

Subset
Only bacterias are present in this cut
```{r}
gfData <- subset_samples(phyloData, region == "GF")
taxGF <- tax_table(gfData)
gfData
```


We need to subset/filter some data.
I follow this tutorial:
https://joey711.github.io/phyloseq/preprocess.html#preprocessing

```{r}
# Filter data based on read count
gfData1 <- filter_taxa(gfData, function(x) sum(x > 3) > (0.2*length(x)), TRUE)

# Remove empty columns and rows, which might have emerged based on the removal of taxa 
gfData2 <- filter_taxa(gfData1, function(x) mean(x) > 0, TRUE) # cols
gfData3 <- prune_samples(sample_sums(gfData2)>0, gfData2) # rows

# Then transform data
gfData4 <- transform_sample_counts(gfData3, function(OTU) OTU / sum(OTU) ) # Transformation of data, so they are relative to the read counts

# Test that the matrix really has no empty cols/rows/any NaNs
#which(rowSums(otu_table(gfData4)) == 0)
#which(colSums(otu_table(gfData4)) == 0)
#sum(is.na(otu_table(gfData4)))

# Then filter taxa based on column mean
GF <- filter_taxa(gfData4, function(x)  mean(x) < 1e-5, TRUE) # Can also be done

GF <- gfData4
```

We end up with 133 samples and 373 taxa!

Graphical summary
```{r}
test <- subset_samples(GF, type == "SW")
```



```{r}
theme_set(theme_bw())
p <- plot_bar(test, "location", fill = "Class") 
p + geom_bar(aes(color = Class, fill = Class), stat = "Identity") #, position = "stack"
```



```{r}
plot_heatmap(test)
```

14
Questions:
- Relative abundance?
- Are we interested in kowing the empty samples?
- Why is GF <- filter_taxa(gfData4, function(x) mean(x) < 1e-5, TRUE) not working?
- the abundance is for the sample numbers. This should be an ASV scale!
