---
title: "2_calculate_threshold"
author: "stinekrye"
date: "2021-11-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
Use the knowledge about the technical replicates to deduce a threshold

Steps:
- Make rarefaction curve to figure out which technical replicates are good enough to keep
- Normalize both samples between 0 and 1
- Make logit transformation
- Compare samples: Calculate mean and stdev
- Make a plot where mean is on the x axis and stddev is on the y axis
- See where the stddeev levels off. Use that mean value to calculate the cut-off from where I do not trust the data anymore. Reverse the logit transformation.

Load data and libraries
```{r}
library(tidyverse)
library(phyloseq)
library(gtools)
library(reshape2)
library(ggpubr)


GF_sampleData <- read.delim("./data/GF_sampleData.txt") #%>% column_to_rownames(., var = "Samples")
GF_asvTable <- read.delim("./data/GF_ASVtable.txt", sep = "\t", check.names = F) #%>% select(., -'row.names')
GF_taxaId <- read.delim("./data/GF_taxaId.txt", sep = "\t") #%>% select(., -'row.names')
```


## Make rarefaction curve


```{r}
# taxo <- GF_taxaId
# taxo$taxonomy <- paste(taxo$Kingdom, taxo$Phylum, taxo$Order, taxo$Family, taxo$Class, taxo$Genus, taxo$V7, sep = ";") # concatenate rows
# taxo <- taxo %>% select(c(taxonomy))                                                                                   # make dataframe ready for merge
# rhea_GF_asvTable <- merge(as.data.frame(GF_asvTable), taxo, by = 0)                                                    # 0 = rownames
# rhea_GF_asvTable <- column_to_rownames(rhea_GF_asvTable, var = "Row.names")                                            # restore rownames
# write.table(rhea_GF_asvTable, file = "GF_ASVtable_RHEA_thresh.txt", sep = "\t", quote = F) # Is moved to the Rhea folder
# head(rhea_GF_asvTable)[124:126]
```


I have changed the RHEA script so the normalized relative abundances sum to 1
```{r, message = FALSE, warning = FALSE}
# source("./code/Rhea-master/1.Normalization/Normalization_get_thresh.R")
# setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21")
```

Read in file with relative abundances
```{r}
GF_asvTable_rel_ab <- read.delim("./code/Rhea-master/1.Normalization/OTUs_Table-norm-rel_thresh.TAB", sep = "\t", check.names = F, row.names = "")
```

They sum to one!
```{r}
total_counts <- as.data.frame(colSums(GF_asvTable_rel_ab[-1]))
```


Figure out which replicates to keep:

```{r}
# Find replicates
GF_sampleData$rowname <-rownames(GF_sampleData)
GF_replicates_sampleData <- GF_sampleData %>% group_by(sampleName) %>% filter(n()>1)
GF_replicates_asvTable <- GF_asvTable_rel_ab[,colnames(GF_asvTable_rel_ab) %in% GF_replicates_sampleData$rowname ] # Some in this list is not in the rel_ab table, so they have been removed

# Find samples with too low sample size
too_low_sample_size <- colnames(GF_asvTable)[colSums(GF_asvTable) < 10000]

# Find the replicates with sufficient sample size
GF_replicates_sampleData <- GF_replicates_sampleData[!GF_replicates_sampleData$rowname %in% too_low_sample_size,]
GF_replicates_asvTable <- GF_replicates_asvTable[,!colnames(GF_replicates_asvTable) %in% too_low_sample_size]


# Remove those which no longer have a duplicate
GF_replicates_sampleData_size_ok <- GF_replicates_sampleData %>% group_by(sampleName) %>% filter(n()>1)
GF_replicates_asvTable_size_ok <- GF_replicates_asvTable %>% select(GF_replicates_sampleData_size_ok$rowname)

# Drop empty rows
GF_replicates_asvTable_rel_size_ok <- GF_replicates_asvTable_size_ok[rowSums(GF_replicates_asvTable_size_ok) > 0,]
```

Produce table with logit trans, sd, mean, which can be used for plotting
```{r}
# Extract identifiers
identifiers <- GF_replicates_sampleData_size_ok %>% select(rowname, sampleID)

# Transpose table so I can merge by columns
table <- as.data.frame(t(GF_replicates_asvTable_rel_size_ok))
table <- merge(table, identifiers, by.x = 0, by.y = "rowname", all.x = TRUE)

# Clean up tabæle, so it only contains ASV abundance values and sample ID
table <- table %>% select(-sampleName) # drop column
table <- table %>%
  select(Row.names, sampleID, everything()) #Rearrange columns
table <- column_to_rownames(table, var = "Row.names")
table <- as.data.frame(t(table)) # Transpose so asvs are rows and samples are columns



# For each col in table make a new col called logit(name)
length = ncol(table)
for (i in 1:length) #
  {
  # Get names + identifier
  name <- colnames(table[i])
  identifier <- table[1,i]
  
  # Do the logit calculation
  # m <- lapply(table[2:nrow(table),i], function(x) x+0.01) # DEAL WITH THIS
  m <- lapply(table[2:nrow(table),i], function(x) log(((x+0.000001-0)/(1-0))/(1-(x-0)/(1-0)))) #+0.000001
  
  # Merge identifiers and m, and ensure the right orientation
  m <- t(as.data.frame(c(identifier,m)))
  
  # Assign the right name to m
  assign(paste("logit_", name, sep = ""),colnames(m))
  
  #Bind m to the table
  table <- cbind(table,m)
}
```


```{r}

# Calculate the mean and std for each pairs of duplicates
# First identify duplicates
logit_table <- as.data.frame(t(table[17:32]))

# Prepare df to dcast by pivoting it longer
test <- gather(logit_table, -sampleID, key = "ASV", value = "Abundance")

# Calculate mean and sd using dcast + combine the dataframes
means <- test %>% dcast(., sampleID ~ ASV, value.var = "Abundance", mean)
sds <- test %>% dcast(., sampleID ~ ASV, value.var = "Abundance", sd)
means$type <- "mean"
sds$type <- "sd"
data <- rbind(means,sds)


# Prepare data for plotting. I want one column with sd and another with mean
data1 <- data %>% pivot_longer(cols = starts_with("ASV"))
data2 <- data1 %>% pivot_wider(names_from = "type", values_from = "value")


# Remove infinite values arising from zeroes
# test <- data2[(data2$mean > -10),]
```


```{r}
data2 %>% ggplot(mapping = aes(x = mean, y = sd))+ # V1 is mean and V2 is std dev
  geom_point(aes(color = factor(sampleID)))+
  geom_smooth(method = lm)+
  stat_cor(aes(color = sampleID), method = "pearson", label.x = -5)+
  xlab("Mean")+
  ylab("Standard deviation")+
  guides(color=guide_legend(title="Sample"))+
  theme_classic()
```


```{r}

data2$sampleID <- factor(data2$sampleID)

data2 %>% ggscatter(x = "mean", y = "sd",
          add = "reg.line",                                 # Add regression line
          # conf.int = TRUE,                                  # Add confidence interval
          color = "sampleID", palette = "jco",           # Color by groups "cyl"
          # shape = "sampleID"
          )+
  stat_cor(aes(color = sampleID), method = "pearson", label.x = -3)  # Add correlation coefficient
```




```{r}
inv.logit(-8)
```

