---
title: "2_data_cleanup"
author: "stinekrye"
date: "2021-09-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
Note: In order to use this script the excel file and R file must share one column called "sample_name", which carries the unique identifier for each data point.
The purpose of this script is:

  1. Rename the rows of taxa.species and columns of seqtab.nochim, so they have consistent ASV# names
  2. Merge the excel file and R file, so they both carries the sort identifier ini_month_date "FG.july.141"
  3. Cut out the part of the data which has the location = GF.
  This script won't save any files unless the # is removed before the save commands
  

```{r}
library(readxl)
library(tidyverse)
```


Import data

```{r}
excel <- read_excel("./data/1_sampleData.xlsx", sheet = "link to R data")
excel <- subset(excel, select = c("Name in R file","ini_month_id"))
rdata <- load("./data/1_seqtab.nochim.OverlapRemoved.rds")
taxa <- load("./data/1_taxa.species.rds")
```
# 1 Rename to ASV# names
## Check if the order of sequences is consistent

```{r}
rownames(taxa.species)[1000] == colnames(seqtab.nochim)[1000] # Yes it is
``` 
## Rename

```{r}
rownames(taxa.species) <- paste("ASV",seq(1:nrow(taxa.species)),sep="")
colnames(seqtab.nochim) <- paste("ASV",seq(1:nrow(taxa.species)),sep="")
```

# 2 Merge
# Merge and save new R file

```{r}
r <- as.data.frame(seqtab.nochim)
r1 <- rownames_to_column(r, var = "sample_name")
test <- merge(r1, excel, by.x = "sample_name", by.y = "Name in R file")
test2 <- column_to_rownames(test, "ini_month_id")
seqtab.nochim.identifier <- test2
#save(seqtab.nochim.identifier, file = "seqtab.nochim.identifier.Rdata" )
```

# 3 Cut out the part of the data with location = GF

## Load data
```{r}
Rdata <- seqtab.nochim.identifier
ExcelGF <- read_excel("./data/1_sampleData.xlsx", sheet = "GF data")
```

# Cut data
```{r}
cut <- seqtab.nochim.identifier[rownames(seqtab.nochim.identifier) %in% ExcelGF$ini_month_id,]
test1 <- rownames(cut) # length = 138
test2 <- ExcelGF$ini_month_id # length = 139. This is one longer than test 1. But why?
GFcutR <- cut
#save(GFcutR, file = "GFcutR.Rdata")
```

# Missing datapoint
The missing datapoint is
```{r}
test2[!test2 %in% test1] # Find what is in test2, which is not in test1. (What is inside the brackets returns a boolean vector which is true if a datapoint in test2 IS NOT in test 1. Thus we cut out the missing datapoint from the test 2 matrix.)
```

