---
title: "2_alpha_diversity"
author: "stinekrye"
date: "2021-11-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Calculate alpha diversity from the clean data

```{r}

```


## 2.3.3.1 Calculate Alpha diversity

### Follow the Rhea workflow
Use the normalized read counts

BE AWARE TO CHANGE THE CUT-OFF VALUES ACCORDINGLY! By default it is 0.5
It also uses the normalized table with empty rows. But it can deal with that
```{r, message = FALSE, warning = FALSE}
source("./code/Rhea-master/2.Alpha-Diversity/Alpha-Diversity.R")
setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21")
```

### Visualization of alpha diversity
Idea: 
Make two plots:
- One with qPCR abundance values against both of the effective indexes (color coded)
- One with one of the effective indexes against the sample name. The point has to be colored in accordance to some grouping Eg. (1st grouping: in situ + sea water, 2nd grouping: Deep, sediment, surface)

  
Merge the sampleData and the alpha-diversity measures in one table to make plotting easier
```{r}
alpha <- read.delim("./code/Rhea-master/2.Alpha-Diversity/alpha-diversity.TAB", sep = "\t", check.names = F, row.names = 1)
alphaSampleData <- merge(sampleData, alpha, by = 0)
```


```{r}
alpha %>% ggplot(mapping = aes(x = Effective.Richness)) +
  geom_histogram(binwidth = 3) +
  theme_classic()
```

```{r}
alpha %>% ggplot(mapping = aes(x = Simpson.Effective, y = Shannon.Effective)) +
  geom_point()
```

