---
title: "5_step_1"
author: "stinekrye"
date: "2021-10-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

The purpose of this script is to illustrate the workflow of how to produce/explore the following:
- 2.3.2.1: Remove singletons, unwanted sequences and negative controls
- 2.3.2.2: Normalize data
        - Rarefaction curves
- 2.3.2.3: Filtering of sequences and samples
- 2.3.3: Alpha diversity  (also include pcr data. correlation of index and pcr abundance), (bacteria, chloroplast).



### Load data and libraries

```{r}
library(tidyverse)
library(readxl)
library(phyloseq)
library(vegan)


sampleData <- read_excel("../data/sampleData.xlsx") %>% column_to_rownames(., var = "Samples")
asvTable <- read.delim("../data/oil_ASVtable.txt", sep = "\t", row.names = "ASVs", check.names = F) #%>% select(., -'row.names')
taxaId <- read.delim("../data/oil_ASVtaxid.txt", sep = "\t", row.names = "ASVs") %>% select(., -'row.names')
```

### Subset data to include only samples from GF
This is done to make the workflow / troubleshooting easier

Create phyloseq object, which makes subsetting all three files easy. Then subset.
```{r}
asvTable <- otu_table(asvTable, taxa_are_rows = TRUE)
sampleData <- sample_data(sampleData)
taxaId <- tax_table(as.matrix(taxaId))
phyloData <- phyloseq(asvTable,sampleData, taxaId)


gfData <- subset_samples(phyloData, region == "GF")


GF_asvTable <- as.data.frame(otu_table(gfData))
GF_sampleData <- as.data.frame(sample_data(gfData))
GF_taxaId <- as.data.frame(tax_table(taxaId))
```

The table needed for Rhea is a tab seperated OTU/ASV table with ASVs as rownames and samplenames as columns. In addidtion to that, we need the taxonomy levels seperated with semicolons as the last column names "taxonomy.
I need to create the table. oil_ASVtable is already a tab seperated file with the right orientation. I need to add the taxonomy column
```{r}
taxo <- GF_taxaId
taxo$taxonomy <- paste(taxo$Kingdom, taxo$Phylum, taxo$Order, taxo$Family, taxo$Class, taxo$Genus, taxo$V7, sep = ";") # concatenate rows
taxo <- taxo %>% select(c(taxonomy))                                                                                   # make dataframe ready for merge
rhea_GF_asvTable <- merge(as.data.frame(GF_asvTable), taxo, by = 0)                                                    # 0 = rownames
rhea_GF_asvTable <- column_to_rownames(rhea_GF_asvTable, var = "Row.names")                                            # restore rownames
#write.table(rhea_GF_asvTable, file = "oil_ASVtable_RHEA.txt", sep = "\t", quote = F) # Is moved to the Rhea folder
```



## 2.3.2.1 Remove singletons, unwanted sequences and negative controls
First we need to remove singletons, chloroplast and mitochondrial species, and negatibe controls.
- Ioannis have removed singletons, chloroplast and mitocondrial species
- We have no negative controls in the in situ samples. By negative controls I mean a control which has not been in contact with any bacteria, but has been sequences to investigate the amount of "background noise" from the project.


## 2.3.2.2 Normalization and removal of samples with low count 

### Rhea Workflow
Run two times. First time to identify underrepresented samples. Which has to be removed
Second time the analysis is made based on the updated OTU table.

The following has been changed in the Rhea Normalization.R file:
setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21/code/Rhea-master/1.Normalization")
file_name <- "oil_ASVtable_RHEA.txt"
labelCutoff <- 10 

```{r}
#source("./code/Rhea-master/1.Normalization/Normalization.R")
#setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21")
```

Use the slope of the rarefaction curve to decide cut-off
- How do you do this?
- My idea:

```{r}
# Exclude samples based on rarefaction curve
rslope <- read.delim("../code/Rhea-master/1.Normalization/1_RarefactionCurve.TAB", sep = "\t", check.names = F)

rslope %>% ggplot(mapping = aes(x = reorder(SampleID, -slope), y = slope)) +
  geom_point() +
  ylim(0, 2)

cut <- rslope[rslope$slope > 0.75,] # set threshold set by visual inspection of the plot

```

```{r}
# Cut away samples based on too low count.
too_low_sample_size <- colnames(GF_asvTable)[colSums(GF_asvTable) < 10000]
```


```{r}
# Cut away unwanted sequences
unwanted <- c(cut$SampleID, too_low_sample_size)
wanted <- colnames(GF_asvTable)[!colnames(GF_asvTable) %in% unwanted]
gfDataCut <- prune_samples(wanted,gfData)
```

Make new Rhea file

```{r}
taxo <- as.data.frame(tax_table(taxaId))
taxo$taxonomy <- paste(taxo$Kingdom, taxo$Phylum, taxo$Order, taxo$Family, taxo$Class, taxo$Genus, taxo$V7, sep = ";") # concatenate rows
taxo <- taxo %>% select(c(taxonomy))                                                                                   # make dataframe ready for merge
rhea_GF_asvTable <- merge(as.data.frame(otu_table(gfDataCut)), taxo, by = 0)                                                    # 0 = rownames
rhea_GF_asvTable <- column_to_rownames(rhea_GF_asvTable, var = "Row.names")                                            # restore rownames
write.table(rhea_GF_asvTable, file = "oil_ASVtable_RHEA_CUT.txt", sep = "\t", quote = F) # Is moved to the Rhea folder
```

Redo Rhea analysis:

The following has been changed in the Rhea Normalization.R file:
setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21/code/Rhea-master/1.Normalization")
file_name <- "oil_ASVtable_RHEA_CUT.txt"
labelCutoff <- 5 

```{r}
#source("../code/Rhea-master/1.Normalization/Normalization.R")
#setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21")
```




## 2.3.3.1 Calculate Alpha diversity

Follow the Rhea workflow

```{r}
#source("../code/Rhea-master/2.Alpha-Diversity/Alpha-Diversity.R")
#setwd("C:/Users/ASUS/Google Drev/AU/Aktuelle/PiB/Project_in_Bioinformatics_E21")
```

Idea: merge the sampleData ans the alpha measure.
Make two plots:
- One with qPCR abundance values against both of the effective indexes (color coded)
- One with one of the effective indexes against the sample name. The point has to be colored in accordance to some grouping Eg. (1st grouping: in situ + sea water, 2nd grouping: Deep, sediment, surface)

```{r}
alpha <- read.delim("../code/Rhea-master/2.Alpha-Diversity/alpha-diversity.TAB", sep = "\t", check.names = F, row.names = 1)
alphaSampleData <- merge(sampleData, alpha, by = 0)
```


```{r}
alphaSampleData %>% ggplot(mapping = aes(x = location, y = Simpson.Effective)) +
  geom_violin() +
  facet_wrap("type")

alphaSampleData %>% ggplot(mapping = aes(x = location, y = Simpson.Effective)) +
  geom_jitter(aes(color = oil)) +
  facet_wrap("type")

alphaSampleData %>% ggplot(mapping = aes(x = location, y = Simpson.Effective)) +
  geom_jitter(aes(color = Evenness)) +
  facet_wrap("type")
```
? How can it make sense that the blanks has high alpha diversity, when Shannon and Simpson index is used?
When using effective richness (enumeration) the distribution looks random.

Notes on Simpson: Look up the formula. Takes the number of species and their abundance into account. A sample with large evenness is considered more diverse than a sample dominated by a few species (possible enrichment)

Notes on Shannon: H = (Relative proportion of species * the natural logarithmic function of that relative proportion) summed over all species. Evenness is H / ln(S) where S is the total number of species in that sample. Low evenness means that a few species dominate




Visualize correlation between QPCR copies and simpson effective numbers. 
? Is it bad to have a correlation? Could that mean a correlation between the size of the sample and the richness? If we have no correlation, does that mean that the effective richness does not depend so much on sample size?
```{r}
alphaSampleData %>% ggplot(mapping = aes(y = QPCRCopies)) + 
  #geom_point(aes(x = Shannon.Effective, color = "Shannon")) +
  geom_point(aes(x = Simpson.Effective, color = "Simpson")) +
  scale_y_log10() +
  #xlim(0,100) +
  NULL
```




Then clean up, prepare presentation and write/clean up introduction. Consider the difference between phyloseq and Rhea
